---
title: "Practice SQL with {sql} Chunks"
author: "Martin Schedlbauer"
output:
  html_document:
    df_print: paged
---

```{r}
library(RSQLite)

# the path below is a Windows path specification
# on MacOS there is no c:/
fpath = "/Users/pengqing/"
dbfile = "orgdb.db"

dbcon <- dbConnect(RSQLite::SQLite(), paste0(fpath,dbfile))
```

Turn on support for foreign key constraint checking

```{sql connection=dbcon}
PRAGMA foreign_keys = ON
```

```{sql connection=dbcon}
select * from Worker
```

```{sql connection=dbcon, output.var="workers.df"}
select * from Worker
```

```{r}
print(workers.df)
```

```{r}
sqlcmd <- "select * from Worker"
workers.df <- dbGetQuery(con = dbcon, sqlcmd)
print(workers.df)
```

```{r}
dbListTables(dbcon)
dbListFields(dbcon, "Worker")
```

Retrieve the last names of all workers under the new column name LastName.
```{sql connection=dbcon}
SELECT LAST_NAME AS LastName 
FROM Worker;
```
Retrieve the first names of all workers in upper case.
```{sql connection=dbcon}
SELECT UPPER(FIRST_NAME)
FROM Worker;
```
Retrieve the names of all departments without duplicates.
```{sql connection=dbcon}
SELECT DISTINCT DEPARTMENT 
FROM Worker;
```

Retrieve the first three characters of all last names of all workers.
```{sql connection=dbcon}
SELECT substr(LAST_NAME, 1, 3) 
FROM Worker;
```
Which worker's last names end with the character 'a'?
```{sql connection=dbcon}
SELECT *
FROM Worker
WHERE LAST_NAME LIKE '%a';
```
Retrieve the first names of all workers and remove any potential leading whitespace (Hint: use rtrim()).
```{sql connection=dbcon}
SELECT rtrim(FIRST_NAME)
FROM Worker;
```
Assume that some department names have trailing whitespace. Retrieve the unique names of all departments with any trailing whitespace removed.
```{sql connection=dbcon}
SELECT DISTINCT rtrim(DEPARTMENT)
FROM Worker;
```
Write an SQL query to list the FIRST_NAME and LAST_NAME from the Worker table into a single column COMPLETE_NAME in the form last, first, e.g., Patel, Vivek. (Hint: SQLite does not support the concat() function; instead it uses the || operator).
```{sql connection=dbcon}
SELECT (LAST_NAME || ',' || FIRST_NAME) AS COMPLETE_NAME
FROM Worker;
```
Retrieve all Worker information sorted by LAST_NAME in ascending order.
```{sql connection=dbcon}
SELECT *
FROM Worker
ORDER BY LAST_NAME ASC;
```
Retrieve the details for all workers with the first name of “Vipul” or “Satish”.
```{sql connection=dbcon}
SELECT *
FROM Worker
WHERE FIRST_NAME = 'Vipul'
OR FIRST_NAME = 'Satish';
```
Retrieve the details for all workers that do not have the first name of either “Vipul” or “Satish”.
```{sql connection=dbcon}
SELECT *
FROM Worker
WHERE FIRST_NAME IS NULL
OR FIRST_NAME = 'Satish'
OR FIRST_NAME = 'Vipul';
```
Which workers have a salary between 100,000 and 500,000? List their last name, first name, and salary in thousands (e.g., 100 rather than 100,000), in that order.
```{sql connection=dbcon}
SELECT LAST_NAME, FIRST_NAME, SALARY/1000 AS SALARY
FROM Worker
WHERE
SALARY BETWEEN 100000 AND 500000;
```
Which workers have a last name that starts with 'S'?
```{sql connection=dbcon}
SELECT * FROM Worker
WHERE
LAST_NAME LIKE 'S%';
```
Which employees joined the company in February 2014? Note that SQLite does not support the year() and month() SQL functions.
```{sql connection=dbcon}
SELECT *
FROM Worker
WHERE JOINING_DATE LIKE '14-02%';
```
How many employees work in the department ‘Admin’?
```{sql connection=dbcon}
SELECT COUNT(*)
FROM Worker
WHERE DEPARTMENT = 'Admin';
```
How many workers work for each department? List the department and number of workers in the descending order of numbers.
```{sql connection=dbcon}
SELECT DEPARTMENT, COUNT(*) AS COUNT
FROM Worker
GROUP BY DEPARTMENT
ORDER BY COUNT DESC;
```
List the first name of all workers who are managers.
```{sql connection=dbcon}
SELECT FIRST_NAME
FROM Worker
WHERE DEPARTMENT = 'Admin';
```
Retrieve the first ten records (rows) from the Worker table sorted by salary from highest to smallest.
```{sql connection=dbcon}
SELECT *
FROM Worker
ORDER BY SALARY DESC
LIMIT 10;
```
Which employees have the same salary? List their first name and salary.
```{sql connection=dbcon}
SELECT a.FIRST_NAME, a.SALARY, b.FIRST_NAME, b.SALARY
FROM Worker a
JOIN Worker b
ON a.SALARY = b.SALARY
AND a.FIRST_NAME != b.FIRST_NAME;
```
Which departments that have fewer than five people in it?
```{sql connection=dbcon}
SELECT DEPARTMENT, COUNT(WORKER_ID) AS 'Number of Workers'
FROM Worker
GROUP BY DEPARTMENT
HAVING COUNT(WORKER_ID) < 5;
```
Which employees have the highest salary in each department?
```{sql connection=dbcon}
SELECT *
FROM (
SELECT max(SALARY) AS TotalSalary, DEPARTMENT 
FROM Worker
GROUP BY DEPARTMENT
) 
AS TempNew
INNER JOIN WORKER t
ON TempNew.DEPARTMENT = t.DEPARTMENT
AND TempNew.TotalSalary = t.Salary;
```
What is the total salary paid out to all employees for each department?
```{sql connection=dbcon}
SELECT DEPARTMENT, SUM(SALARY) AS SUM
FROM Worker
GROUP BY DEPARTMENT;
```
Who are the employees who earn the highest salary?
```{sql connection=dbcon}
SELECT *
FROM Worker
WHERE SALARY = (
SELECT max(SALARY)
FROM Worker
);
```


```{r}
dbDisconnect(dbcon)
```

