---
title: "R Notebook"
output: html_notebook
---

```{r}
#install.packages("RMySQL")
library(RMySQL)
library(dplyr)

mydb<-dbConnect(MySQL(), user='root', password='0608Mysql!', dbname='practicum', host='localhost')
```

```{r}
dbListTables(mydb)
```

```{r}
sql <- "DROP TABLE IF EXISTS `aircraft_model`;"
dbExecute(mydb, sql)
```
```{r}
sql <- "CREATE TABLE `aircraft_model` (
  `amid` int NOT NULL AUTO_INCREMENT,
  `model` varchar(100) DEFAULT NULL,
  PRIMARY KEY (`amid`)
);"
dbExecute(mydb, sql)
```

```{r}
sql <- "DROP TABLE IF EXISTS `aircraft_type`;"
dbExecute(mydb, sql)
```
```{r}
sql <- "CREATE TABLE `aircraft_type` (
  `atid` int NOT NULL AUTO_INCREMENT,
  `type` varchar(45) DEFAULT NULL,
  PRIMARY KEY (`atid`),
  UNIQUE KEY `type_UNIQUE` (`type`)
);"
dbExecute(mydb, sql)
```

```{r}
sql <- "DROP TABLE IF EXISTS `aircraft`;"
dbExecute(mydb, sql)
```

```{r}
sql <- "CREATE TABLE `aircraft` (
  `acid` int NOT NULL AUTO_INCREMENT,
  `amid` int DEFAULT NULL,
  `number_of_engines` varchar(10) DEFAULT NULL,
  `is_large` tinyint DEFAULT NULL,
  `atid` int DEFAULT NULL,
  PRIMARY KEY (`acid`),
  KEY `amid_idx` (`amid`),
  KEY `atid_idx` (`atid`),
  CONSTRAINT `amid` FOREIGN KEY (`amid`) REFERENCES `aircraft_model` (`amid`),
  CONSTRAINT `atid` FOREIGN KEY (`atid`) REFERENCES `aircraft_type` (`atid`)
);"
dbExecute(mydb, sql)
```
```{r}
sql <- "DROP TABLE IF EXISTS `airline`;"
dbExecute(mydb, sql)
```
```{r}
sql <- "CREATE TABLE `airline` (
  `alid` int NOT NULL AUTO_INCREMENT,
  `name` varchar(200) DEFAULT NULL,
  PRIMARY KEY (`alid`)
);"
dbExecute(mydb, sql)
```
```{r}
sql <- "DROP TABLE IF EXISTS `airport`;"
dbExecute(mydb, sql)
```
```{r}
sql <- "CREATE TABLE `airport` (
  `apid` int NOT NULL AUTO_INCREMENT,
  `name` varchar(45) DEFAULT NULL,
  `code` varchar(45) DEFAULT NULL,
  `state` varchar(45) DEFAULT NULL,
  PRIMARY KEY (`apid`)
);"
dbExecute(mydb, sql)
```
```{r}
sql <- "DROP TABLE IF EXISTS `condition`;"
dbExecute(mydb, sql)
```
```{r}
sql <- "CREATE TABLE `condition` (
  `cid` int NOT NULL AUTO_INCREMENT,
  `sky_condition` varchar(45) DEFAULT NULL,
  `explanation` varchar(2000) DEFAULT NULL,
  PRIMARY KEY (`cid`)
);"
dbExecute(mydb, sql)
```
```{r}
sql <- "DROP TABLE IF EXISTS `incident`;"
dbExecute(mydb, sql)
```
```{r}
sql <- "CREATE TABLE `incident` (
  `iid` int NOT NULL AUTO_INCREMENT,
  `date` datetime DEFAULT NULL,
  `origin` varchar(45) DEFAULT NULL,
  `alid` int DEFAULT NULL,
  `acid` int DEFAULT NULL,
  `impact` tinyint DEFAULT NULL,
  `cid` int DEFAULT NULL,
  `apid` int DEFAULT NULL,
  `flightPhase` enum('takeoff','inflight','unkown','landing') DEFAULT NULL,
  PRIMARY KEY (`iid`),
  KEY `alid_idx` (`alid`),
  KEY `acid_idx` (`acid`),
  KEY `cid_idx` (`cid`),
  KEY `apid_idx` (`apid`),
  CONSTRAINT `acid` FOREIGN KEY (`acid`) REFERENCES `aircraft` (`acid`),
  CONSTRAINT `alid` FOREIGN KEY (`alid`) REFERENCES `airline` (`alid`),
  CONSTRAINT `apid` FOREIGN KEY (`apid`) REFERENCES `airport` (`apid`),
  CONSTRAINT `cid` FOREIGN KEY (`cid`) REFERENCES `condition` (`cid`)
);"
dbExecute(mydb, sql)
```
```{r}
dbListTables(mydb)
```
```{r}
fn<-'BirdStrikesData.csv'

df.raw<-read.csv(file=fn,
                 header=T,
                 stringsAsFactors = F)
```

```{r}
head(df.raw)
```
```{r}
library(sqldf)
```
Load data into table aircraft_model
```{r}
detach("package:RMySQL", unload=TRUE)
df.aircraft.model <- sqldf("SELECT `Aircraft..Make.Model` AS model FROM `df.raw` GROUP BY model")
```
```{r}
library(RMySQL)
dbWriteTable(mydb, "aircraft_model", df.aircraft.model, append=TRUE, row.names=FALSE)
```
Assign df.aircraft.model to the data frame of table aircraft_model
```{r}
df.aircraft.model <- data.frame(tbl(mydb, "aircraft_model"))
```

Load data into table aircraft_type
```{r}
detach("package:RMySQL", unload=TRUE)
df.aircraft.type <- sqldf("SELECT `Aircraft..Type` AS type FROM `df.raw` GROUP BY type")
```
```{r}
library(RMySQL)
dbWriteTable(mydb, "aircraft_type", df.aircraft.type, append=TRUE, row.names=FALSE)
df.aircraft.type <- data.frame(tbl(mydb, "aircraft_type"), TRUE, TRUE)
```
Load data into table aircraft

```{r}
df.aircraft <- data.frame(tbl(mydb, "aircraft"))
```


```{r}
n.df.raw <- nrow(df.raw)
for (r in 1:n.df.raw) {
  amid <- df.aircraft.model$amid[which(
    df.raw$Aircraft..Make.Model[r] == df.aircraft.model$model)]
  atid <- df.aircraft.type$atid[which(
    df.raw$Aircraft..Type[r] == df.aircraft.type$type)]
  if (df.raw$Is.Aircraft.Large.[r] == "yes") {
    large <- 1;
  } else {
    large <- 0;
  }
  number <- df.raw$Aircraft..Number.of.engines.[r]
  if (is.null(number) | length(number) == 0 | number == "" | number == 'C') {
    next
  }
  command <- paste0("INSERT INTO aircraft(amid, atid, number_of_engines, is_large) VALUES (",
                          amid,",", atid,",", number,",", large, ");")
  #print(command)
  dbExecute(mydb, command)
}

```

```{r}
df.aircraft <- data.frame(tbl(mydb, "aircraft"))
```
```{r}
dbReadTable(mydb, "aircraft")
```

Load data into table airline
```{r}
detach("package:RMySQL", unload=TRUE)
df.airline <- sqldf("SELECT `Aircraft..Airline.Operator` AS name FROM `df.raw` GROUP BY name")
```
```{r}
library(RMySQL)
dbWriteTable(mydb, "airline", df.airline, append=TRUE, row.names=FALSE)
df.airline <- data.frame(tbl(mydb, "airline"))
```
Load data into table airport
```{r}
detach("package:RMySQL", unload=TRUE)
df.airport <- sqldf("SELECT `Airport..Name` AS name, `Origin.State` AS state FROM `df.raw` GROUP BY name")
```
```{r}
library(RMySQL)
dbWriteTable(mydb, "airport", df.airport, append=TRUE, row.names=FALSE)
df.airport <- data.frame(tbl(mydb, "airport"))
```
Load data into table condition
```{r}
detach("package:RMySQL", unload=TRUE)
df.condition <- sqldf("SELECT `Conditions..Sky` AS sky_condition FROM `df.raw` GROUP BY sky_condition")
```
```{r}
library(RMySQL)
dbWriteTable(mydb, "condition", df.condition, append=TRUE, row.names=FALSE)
df.condition <- data.frame(tbl(mydb, "condition"))
```
Load data into table incident
```{r}
df.incident <- data.frame(tbl(mydb, "incident"))
```
  amid <- df.aircraft.model$amid[which(
    df.raw$Aircraft..Make.Model[r] == df.aircraft.model$model)]
  atid <- df.aircraft.type$atid[which(
    df.raw$Aircraft..Type[r] == df.aircraft.type$type)]
  if (df.raw$Is.Aircraft.Large.[r] == "yes") {
    large <- 1;
  } else {
    large <- 0;
  }
  number <- df.raw$Aircraft..Number.of.engines.[r]
```{r}
n.df.raw <- nrow(df.raw)
for (r in 1:n.df.raw) {
  date <- df.raw$FlightDate[r]
  
  if (!is.null(date) && date != '') {
    date <- format(as.POSIXct(date,format='%m/%d/%Y %H:%M'),format='%m/%d/%Y')
    date <- as.Date(date, format = "%m/%d/%Y")
  } else{
    next
  }
  origin <- df.raw$Origin.State[r]
  alid <- df.airline$alid[which(
    df.raw$Aircraft..Airline.Operator[r] == df.airline$name)]
  amid <- df.aircraft.model$amid[which(
    df.raw$Aircraft..Make.Model[r] == df.aircraft.model$model)]
  
  n.df.ac <- nrow(df.aircraft)
  for (i in 1:n.df.ac) {
    if (df.aircraft$amid[i] == amid) {
      acid <- df.aircraft$acid[i]
      break
    }
  }

  flightPhase <- df.raw$When..Phase.of.flight[r]
  if(flightPhase == "Landing Roll" || flightPhase == "Approach"){
    flightPhase = "landing"
  }else if(flightPhase == "Climb" || flightPhase == "Take-off run"){
    flightPhase = "takeoff"
  } else if(flightPhase == "Descent" || flightPhase == "Taxi"){
    flightPhase = "inflight"
  } else{
    flightPhase = "unkown"
  }
  
  
  impact <- df.raw$Effect..Impact.to.flight[r]
  if (impact == "None") {
    impact = 0
  } else {
    impact = 1
  }
  cid <- df.condition$cid[which(
    df.raw$Conditions..Sky[r] == df.condition$sky_condition
  )]
  apid <- df.airport$apid[which(
    df.raw$Airport..Name[r] == df.airport$name
  )]
  if (is.null(acid) | length(acid) == 0) {
    print("yes")
    next
  }
  command <- paste0("INSERT INTO incident(date,origin,alid,acid,impact,cid,apid,flightPhase) VALUES (\"",
                    date,"\",\"",origin,"\",\"",alid,"\",\"",acid,"\",\"",impact,"\",\"",cid,"\",\"",apid,"\",\"",flightPhase,"\");")

  #print(command)
  dbExecute(mydb, command)
}

```


```{r}
# construct a SQL query
sqlCmd = "SELECT * FROM incident"

# send the SQL query to the database
rs <- dbGetQuery(mydb, sqlCmd)
head(rs)

```
```{r}
# construct a SQL query
sqlCmd = "SELECT * FROM airport"

# send the SQL query to the database
rs <- dbGetQuery(mydb, sqlCmd)
head(rs)

```

```{r}
dbListTables(mydb)
```

```{r}
# construct a SQL query
sqlCmd = "SELECT * FROM condition"

# send the SQL query to the database
rs <- dbGetQuery(mydb, sqlCmd)
head(rs)

```
```{r}
# 4
cmd = "select flightPhase, count(*) as Count
from incident
group by flightPhase"

# send the SQL query to the database
rs <- dbGetQuery(mydb, cmd)
show(rs)

```

```{r}
# 5
sqlCmd = "with countTable as (select flightPhase as flightPhase, count(*) as numIncident
from incident
group by flightPhase)
select flightPhase 
from countTable 
where numIncident > (select AVG(numIncident) from countTable);"

# send the SQL query to the database
rs <- dbGetQuery(mydb, sqlCmd)
show(rs)

```

```{r}
#6
sqlcmd = "SELECT Month(date) AS incidentMonth, ROUND((COUNT(*) / (select MAX(Year(date)) - MIN(Year(date)) + 1 from incident)), 2) AS Count
FROM incident
GROUP BY incidentMonth 
ORDER BY incidentMonth;"

rs <- dbGetQuery(mydb, sqlcmd)
show(rs)
```

```{r}
#7
#install.packages("tidyverse")
library(ggplot2)
sql <- "SELECT YEAR(date) AS Year, COUNT(*) AS Count FROM incident WHERE date BETWEEN '2000-1-1' AND '2011-12-31' GROUP BY YEAR(date)"
df.incident.year <- data.frame(tbl(mydb, sql(sql)))
chart <- ggplot(data=df.incident.year, aes(x=Year, y=Count)) +
    geom_bar(stat="identity") +
    ggtitle("Average Amount of Incidents per Year") +
    labs(x="Year", y="Amount of Incidents") +
    theme_bw() +
    theme(title=element_text(size=16, color="blue3"))
print(chart)
```
```{r}
CREATE procedure InsertIncident 
  @
AS
BEGIN
  
END
GO
```


```{r}
sql <- "DELETE FROM incident";
dbExecute(mydb, sql)
```
```{r}
sql <- "DELETE FROM airline";
dbExecute(mydb, sql)
```
```{r}
dbDisconnect(mydb)
```


